---
title: "HRS Stress Dataset Construction"
subtitle: "CHORDS Lab – Washington State University"
author:
- "Jason Cross"
- "Shawna Beese"
date: "Last updated: 2025-10-29"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

## Overview

This R Markdown script builds the HRS Stress Dataset for allostatic (over)load research using self-reported stress measures from the [Health and Retirement Study (HRS)](https://hrs.isr.umich.edu/).  

⚠️ Before continuing, ensure your data are downloaded and organized as described in [`HRS_stress_access_guide.md`](https://github.com/CHORDSLab/HRS-Stress-Measures-and-Supplementary-Datasets/blob/main/HRS_stress/HRS_stress_access_guide.md).  

For a full overview of variables, see [`HRS_Stress_Data_Overview.pdf`](https://github.com/CHORDSLab/HRS-Stress-Measures-and-Supplementary-Datasets/blob/main/HRS_stress/HRS_Stress_Data_Overview.pdf).  

A user guide for the Psychosocial and Lifestyle Questionnaire (2006–2022) is available here: [HRS 2006–2022 SAQ User Guide (PDF)](https://hrs.isr.umich.edu/sites/default/files/biblio/HRS%202006-2022%20SAQ%20User%20Guide.pdf).  


---

## Build Modes (Read Me First)  

There are two ways to use this script, depending on whether you want to build a **single-year** or **longitudinal** version of the HRS Stress dataset.  

---

### Option A. Single-Year Build  
Use this mode if you only downloaded **one** HRS Core wave (e.g., 2016).  

**Steps:**  
1. Run the following chunks:  
   - `Setup & Libraries`  
   - `Set File Path`  
   - the `Load and clean [YEAR]` chunk for your chosen year    
2. (Optional) Run the “Save [YEAR]-only dataset” chunk right below it.    
3. Stop here — you do **not** need to combine other years.  

💡 *Example:* If you downloaded only the 2016 Core data, scroll down to the “Load and clean 2016” section and run those chunks.  

---

### Option B. Longitudinal Build (2006–2022)  
Use this mode if you downloaded **all HRS Core waves** from 2006–2022.  

**Steps:**  
1. Run every chunk in order, from top to bottom:  
   - `Setup & Libraries`  
   - `Step 1. Set File Path`  
   - all `Step 2. Load and clean [YEAR]` chunks (2006–2022)  
   - `Step 3. Combine All Years (Longitudinal Build Only)`  
   - `Step 4. Save and Export Longitudinal Dataset`  
2. This will create a single combined dataset named `full_hrs_stress`.  

💡 *Partial build:* You can also build a smaller panel (e.g., 2014–2022) by running only the chunks for those specific years, then running the combine step.  


---

### What You’ll Need to Run This Script  

- Raw HRS data, organized as described in the guide  
- R environment with the following packages: `SAScii`, `dplyr`, `stringr`  
- Familiarity with HRS data formats and .sas scripts  


## Setup & Libraries  

```{r install-packages, eval=FALSE}
install.packages(c("SAScii", "dplyr", "stringr"))
```

```{r libraries}
library(SAScii)
library(dplyr)
library(stringr)
```

---

## 📁 Set File Path  

```{r file-path}
# Set to the location of your "HRS Data Products" folder
path <- "../path_to_your_folder/HRS Data Products/HRS Core Data" # Replace with the full path to your local "HRS Data Products" folder
```



---

### Load and clean 2006 HRS Core (LB) data *(run for 2006 builds or longitudinal builds)*

```{r}

#####################
# Load and clean 2006 HRS Core (LBQ) data
# Run this chunk if you need 2006 (either single-year OR longitudinal)
#####################

# Navigate to the HRS Core Data folder
setwd(file.path(path, "h06core"))

# Load 2006 Leave-Behind Questionnaire (LBQ)
core_2006 <- read.SAScii(
  "h06da/H06LB_R.da",
  "h06sas/H06LB_R.sas"
)

# Add unique identifier (HHID + PN)
add_id <- function(df) df %>% mutate(unique_id = paste(HHID, PN, sep = "_"))
core_2006 <- add_id(core_2006)

clean_2006 <- core_2006 %>%
  transmute(
    unique_id,
    survey_year = 2006L,
    id_year = paste(unique_id, survey_year, sep = "_"),

    # End-of-Month Financial Strain
    eomfs   = KLB039B,

    # Ongoing Chronic Stressors (A1–A8)
    ocs_01  = KLB040A,
    ocs_02  = KLB040B,
    ocs_03  = KLB040C,
    ocs_04  = KLB040D,
    ocs_05  = KLB040E,
    ocs_06  = KLB040F,
    ocs_07  = KLB040G,
    ocs_08  = KLB040H,

    # Stressful Life Events (last 5 years) + year follow-ups
    sle_01  = KLB038A,   # job loss
    sle_02  = KLB038AY,  # job loss year
    sle_03  = KLB038B,   # unemployed self
    sle_04  = KLB038BY,  # unemployed self year
    sle_05  = KLB038C,   # other unemployed
    sle_06  = KLB038CY,  # other unemployed year
    sle_07  = KLB038D,   # moved to worse residence
    sle_08  = KLB038DY,  # moved year
    sle_09  = KLB038E,   # robbed/burglarized
    sle_10  = KLB038EY   # robbed/burglarized year
    # Note: 2006 has no fraud items (sle_11 / sle_12)
  )

# Preview the cleaned 2006 data
glimpse(clean_2006)

```

### Optional: Save 2006-only dataset

```{r}
# Run this if you're only building 2006 and not doing longitudinal

write.csv(clean_2006, "HRS_Stress_2006.csv", row.names = FALSE)
saveRDS(clean_2006, "HRS_Stress_2006.rds")

```



---

### Load and clean 2008 HRS Core (LB) data *(run for 2008 builds or longitudinal builds)*

```{r}

#####################
# Load and clean 2008 HRS Core (LBQ) data
# Run this chunk if you need 2008 (either single-year OR longitudinal)
#####################

# Navigate to the HRS Core Data folder
setwd(file.path(path, "h08core"))

# Load 2008 Leave-Behind Questionnaire (LBQ)
core_2008 <- read.SAScii(
  "h08da/H08LB_R.da",
  "h08sas/H08LB_R.sas"
)

# Add unique identifier (HHID + PN)
add_id <- function(df) df %>% mutate(unique_id = paste(HHID, PN, sep = "_"))
core_2008 <- add_id(core_2008)

clean_2008 <- core_2008 %>%
  transmute(
    unique_id,
    survey_year = 2008L,
    id_year = paste(unique_id, survey_year, sep = "_"),

    # End-of-Month Financial Strain
    eomfs   = LLB040,

    # Stressful Life Events (last 5 years) + year follow-ups
    sle_01  = LLB038A,
    sle_02  = LLB038AY,
    sle_03  = LLB038B,
    sle_04  = LLB038BY,
    sle_05  = LLB038C,
    sle_06  = LLB038CY,
    sle_07  = LLB038D,
    sle_08  = LLB038DY,
    sle_09  = LLB038E,
    sle_10  = LLB038EY,
    sle_11  = LLB038F,
    sle_12  = LLB038FY
  )

# Preview the cleaned 2008 data
glimpse(clean_2008)

```

### Optional: Save 2008-only dataset

```{r}
# Run this if you're only building 2008 and not doing longitudinal

write.csv(clean_2008, "HRS_Stress_2008.csv", row.names = FALSE)
saveRDS(clean_2008, "HRS_Stress_2008.rds")

```



---

### Load and clean 2010 HRS Core (LB) data *(run for 2010 builds or longitudinal builds)*

```{r}

#####################
# Load and clean 2010 HRS Core (LBQ) data
# Run this chunk if you need 2010 (either single-year OR longitudinal)
#####################

# Navigate to the HRS Core Data folder
setwd(file.path(path, "h10core"))

# Load 2010 Leave-Behind Questionnaire (LBQ)
core_2010 <- read.SAScii(
  "h10da/H10LB_R.da",
  "h10sas/H10LB_R.sas"
)

# Add unique identifier (HHID + PN)
add_id <- function(df) df %>% mutate(unique_id = paste(HHID, PN, sep = "_"))
core_2010 <- add_id(core_2010)

clean_2010 <- core_2010 %>%
  transmute(
    unique_id,
    survey_year = 2010L,
    id_year = paste(unique_id, survey_year, sep = "_"),

    # End-of-Month Financial Strain
    eomfs  = MLB040,

    # Ongoing Chronic Stressors (A1–A8)
    ocs_01 = MLB040A_A,
    ocs_02 = MLB040A_B,
    ocs_03 = MLB040A_C,
    ocs_04 = MLB040A_D,
    ocs_05 = MLB040A_E,
    ocs_06 = MLB040A_F,
    ocs_07 = MLB040A_G,
    ocs_08 = MLB040A_H,

    # Stressful Life Events (last 5 years) + year follow-ups
    sle_01 = MLB038A,
    sle_02 = MLB038AY,
    sle_03 = MLB038B,
    sle_04 = MLB038BY,
    sle_05 = MLB038C,
    sle_06 = MLB038CY,
    sle_07 = MLB038D,
    sle_08 = MLB038DY,
    sle_09 = MLB038E,
    sle_10 = MLB038EY,
    sle_11 = MLB038F,
    sle_12 = MLB038FY
  )

# Preview the cleaned 2010 data
glimpse(clean_2010)

```

### Optional: Save 2010-only dataset

```{r}
# Run this if you're only building 2010 and not doing longitudinal

write.csv(clean_2010, "HRS_Stress_2010.csv", row.names = FALSE)
saveRDS(clean_2010, "HRS_Stress_2010.rds")

```



---

### Load and clean 2012 HRS Core (LB) data *(run for 2012 builds or longitudinal builds)*

```{r}

#####################
# Load and clean 2012 HRS Core (LBQ) data
# Run this chunk if you need 2012 (either single-year OR longitudinal)
#####################

# Navigate to the HRS Core Data folder
setwd(file.path(path, "h12core"))

# Load 2012 Leave-Behind Questionnaire (LBQ)
core_2012 <- read.SAScii(
  "h12da/H12LB_R.da",
  "h12sas/H12LB_R.sas"
)

# Add unique identifier (HHID + PN)
add_id <- function(df) df %>% mutate(unique_id = paste(HHID, PN, sep = "_"))
core_2012 <- add_id(core_2012)

clean_2012 <- core_2012 %>%
  transmute(
    unique_id,
    survey_year = 2012L,
    id_year = paste(unique_id, survey_year, sep = "_"),

    # End-of-Month Financial Strain
    eomfs  = NLB040,

    # Ongoing Chronic Stressors (A1–A8)
    ocs_01 = NLB040A_A,
    ocs_02 = NLB040A_B,
    ocs_03 = NLB040A_C,
    ocs_04 = NLB040A_D,
    ocs_05 = NLB040A_E,
    ocs_06 = NLB040A_F,
    ocs_07 = NLB040A_G,
    ocs_08 = NLB040A_H,

    # Stressful Life Events (last 5 years) + year follow-ups
    sle_01 = NLB038A,
    sle_02 = NLB038AY,
    sle_03 = NLB038B,
    sle_04 = NLB038BY,
    sle_05 = NLB038C,
    sle_06 = NLB038CY,
    sle_07 = NLB038D,
    sle_08 = NLB038DY,
    sle_09 = NLB038E,
    sle_10 = NLB038EY,
    sle_11 = NLB038F,
    sle_12 = NLB038FY
  )

# Preview the cleaned 2012 data
glimpse(clean_2012)


```

### Optional: Save 2012-only dataset

```{r}
# Run this if you're only building 2012 and not doing longitudinal

write.csv(clean_2012, "HRS_Stress_2012.csv", row.names = FALSE)
saveRDS(clean_2012, "HRS_Stress_2012.rds")

```



---

### Load and clean 2014 HRS Core (LB) data *(run for 2014 builds or longitudinal builds)*

```{r}

#####################
# Load and clean 2014 HRS Core (LBQ) data
# Run this chunk if you need 2014 (either single-year OR longitudinal)
#####################

# Navigate to the HRS Core Data folder
setwd(file.path(path, "h14core"))

# Load 2014 Leave-Behind Questionnaire (LBQ)
core_2014 <- read.SAScii(
  "h14da/H14LB_R.da",
  "h14sas/H14LB_R.sas"
)

# Add unique identifier (HHID + PN)
add_id <- function(df) df %>% mutate(unique_id = paste(HHID, PN, sep = "_"))
core_2014 <- add_id(core_2014)

clean_2014 <- core_2014 %>%
  transmute(
    unique_id,
    survey_year = 2014L,
    id_year = paste(unique_id, survey_year, sep = "_"),

    # End-of-Month Financial Strain
    eomfs  = OLB035,

    # Ongoing Chronic Stressors (A1–A8)
    ocs_01 = OLB035A_1,
    ocs_02 = OLB035A_2,
    ocs_03 = OLB035A_3,
    ocs_04 = OLB035A_4,
    ocs_05 = OLB035A_5,
    ocs_06 = OLB035A_6,
    ocs_07 = OLB035A_7,
    ocs_08 = OLB035A_8
  )

# Preview the cleaned 2014 data
glimpse(clean_2014)


```

### Optional: Save 2014-only dataset

```{r}
# Run this if you're only building 2014 and not doing longitudinal

write.csv(clean_2014, "HRS_Stress_2014.csv", row.names = FALSE)
saveRDS(clean_2014, "HRS_Stress_2014.rds")

```



---

### Load and clean 2016 HRS Core (LB) data *(run for 2016 builds or longitudinal builds)*

```{r}

#####################
# Load and clean 2016 HRS Core (LBQ) data
# Run this chunk if you need 2016 (either single-year OR longitudinal)
#####################

# Navigate to the HRS Core Data folder
setwd(file.path(path, "h16core"))

# Load 2016 Leave-Behind Questionnaire (LBQ)
core_2016 <- read.SAScii(
  "h16da/H16LB_R.da",
  "h16sas/H16LB_R.sas"
)

# Add unique identifier (HHID + PN)
add_id <- function(df) df %>% mutate(unique_id = paste(HHID, PN, sep = "_"))
core_2016 <- add_id(core_2016)

clean_2016 <- core_2016 %>%
  transmute(
    unique_id,
    survey_year = 2016L,
    id_year = paste(unique_id, survey_year, sep = "_"),

    # End-of-Month Financial Strain
    eomfs  = PLB035,

    # Ongoing Chronic Stressors (A1–A8)
    ocs_01 = PLB035A_1,
    ocs_02 = PLB035A_2,
    ocs_03 = PLB035A_3,
    ocs_04 = PLB035A_4,
    ocs_05 = PLB035A_5,
    ocs_06 = PLB035A_6,
    ocs_07 = PLB035A_7,
    ocs_08 = PLB035A_8
  )

# Preview the cleaned 2016 data
glimpse(clean_2016)


```

### Optional: Save 2016-only dataset

```{r}
# Run this if you're only building 2016 and not doing longitudinal

write.csv(clean_2016, "HRS_Stress_2016.csv", row.names = FALSE)
saveRDS(clean_2016, "HRS_Stress_2016.rds")

```



---

### Load and clean 2018 HRS Core (LB) data *(run for 2018 builds or longitudinal builds)*

```{r}

#####################
# Load and clean 2018 HRS Core (LBQ) data
# Run this chunk if you need 2018 (either single-year OR longitudinal)
#####################

# Navigate to the HRS Core Data folder
setwd(file.path(path, "h18core"))

# Load 2018 Leave-Behind Questionnaire (LBQ)
core_2018 <- read.SAScii(
  "h18da/H18LB_R.da",
  "h18sas/H18LB_R.sas"
)

# Add unique identifier (HHID + PN)
add_id <- function(df) df %>% mutate(unique_id = paste(HHID, PN, sep = "_"))
core_2018 <- add_id(core_2018)

clean_2018 <- core_2018 %>%
  transmute(
    unique_id,
    survey_year = 2018L,
    id_year = paste(unique_id, survey_year, sep = "_"),

    # End-of-Month Financial Strain
    eomfs  = QLB035,

    # Ongoing Chronic Stressors (A1–A8)
    ocs_01 = QLB035A_1,
    ocs_02 = QLB035A_2,
    ocs_03 = QLB035A_3,
    ocs_04 = QLB035A_4,
    ocs_05 = QLB035A_5,
    ocs_06 = QLB035A_6,
    ocs_07 = QLB035A_7,
    ocs_08 = QLB035A_8,

    # Perceived Stress Scale (PSS-10)
    # Reminder: 4,5,7,8 are usually reverse-scored
    pss_01 = QLB035B1,
    pss_02 = QLB035B2,
    pss_03 = QLB035B3,
    pss_04 = QLB035B4,  # reverse
    pss_05 = QLB035B5,  # reverse
    pss_06 = QLB035B6,
    pss_07 = QLB035B7,  # reverse
    pss_08 = QLB035B8,  # reverse
    pss_09 = QLB035B9,
    pss_10 = QLB035B10
  )

# Preview the cleaned 2018 data
glimpse(clean_2018)


```

### Optional: Save 2018-only dataset

```{r}
# Run this if you're only building 2018 and not doing longitudinal

write.csv(clean_2018, "HRS_Stress_2018.csv", row.names = FALSE)
saveRDS(clean_2018, "HRS_Stress_2018.rds")

```



---

### Load and clean 2020 HRS Core (LB) data *(run for 2020 builds or longitudinal builds)*

```{r}

#####################
# Load and clean 2020 HRS Core (LBQ) data
# Run this chunk if you need 2020 (either single-year OR longitudinal)
#####################

# Navigate to the HRS Core Data folder
setwd(file.path(path, "h20core"))

# Load 2020 Leave-Behind Questionnaire (LBQ)
core_2020 <- read.SAScii(
  "h20da/H20LB_R.da",
  "h20sas/H20LB_R.sas"
)

# Add unique identifier (HHID + PN)
add_id <- function(df) df %>% mutate(unique_id = paste(HHID, PN, sep = "_"))
core_2020 <- add_id(core_2020)

clean_2020 <- core_2020 %>%
  transmute(
    unique_id,
    survey_year = 2020L,
    id_year = paste(unique_id, survey_year, sep = "_"),

    # End-of-Month Financial Strain
    eomfs   = RLB035,

    # Ongoing Chronic Stressors (A1–A8)
    ocs_01  = RLB035A1,
    ocs_02  = RLB035A2,
    ocs_03  = RLB035A3,
    ocs_04  = RLB035A4,
    ocs_05  = RLB035A5,
    ocs_06  = RLB035A6,
    ocs_07  = RLB035A7,
    ocs_08  = RLB035A8,

    # Perceived Stress Scale (PSS-10)
    # Reminder: items 4,5,7,8 are usually reverse-scored
    pss_01  = RLB035B1,
    pss_02  = RLB035B2,
    pss_03  = RLB035B3,
    pss_04  = RLB035B4,  # reverse
    pss_05  = RLB035B5,  # reverse
    pss_06  = RLB035B6,
    pss_07  = RLB035B7,  # reverse
    pss_08  = RLB035B8,  # reverse
    pss_09  = RLB035B9,
    pss_10  = RLB035B10,

    # Stressful Life Events (last 5 years)
    sle_01  = RLB037B1,
    sle_02  = RLB037B1A,
    sle_03  = RLB037B2,
    sle_04  = RLB037B2A,
    sle_05  = RLB037B3,
    sle_06  = RLB037B3A,
    sle_07  = RLB037B4,
    sle_08  = RLB037B4A,
    sle_09  = RLB037B5,
    sle_10  = RLB037B5A,
    sle_11  = RLB037B6,
    sle_12  = RLB037B6A,

  )

# Preview the cleaned 2020 data
glimpse(clean_2020)


```

### Optional: Save 2020-only dataset

```{r}
# Run this if you're only building 2020 and not doing longitudinal

write.csv(clean_2020, "HRS_Stress_2020.csv", row.names = FALSE)
saveRDS(clean_2020, "HRS_Stress_2020.rds")

```



---

### Load and clean 2022 HRS Core (LB) data *(run for 2022 builds or longitudinal builds)*

```{r}

#####################
# Load and clean 2022 HRS Core (LBQ) data
# Run this chunk if you need 2022 (either single-year OR longitudinal)
#####################

# Navigate to the HRS Core Data folder
setwd(file.path(path, "h22core"))

# Load 2022 Leave-Behind Questionnaire (LBQ)
core_2022 <- read.SAScii(
  "h22da/H22LB_R.da",
  "h22sas/H22LB_R.sas"
)

# Add unique identifier (HHID + PN)
add_id <- function(df) df %>% mutate(unique_id = paste(HHID, PN, sep = "_"))
core_2022 <- add_id(core_2022)

clean_2022 <- core_2022 %>%
  transmute(
    unique_id,
    survey_year = 2022L,
    id_year = paste(unique_id, survey_year, sep = "_"),

    # End-of-Month Financial Strain
    eomfs  = SLB035,

    # Ongoing Chronic Stressors (A1–A8)
    ocs_01 = SLB035A1,
    ocs_02 = SLB035A2,
    ocs_03 = SLB035A3,
    ocs_04 = SLB035A4,
    ocs_05 = SLB035A5,
    ocs_06 = SLB035A6,
    ocs_07 = SLB035A7,
    ocs_08 = SLB035A8,

    # Perceived Stress Scale (PSS-10)
    # Reminder: 4,5,7,8 are usually reverse-scored
    pss_01 = SLB035B1,
    pss_02 = SLB035B2,
    pss_03 = SLB035B3,
    pss_04 = SLB035B4,  # reverse
    pss_05 = SLB035B5,  # reverse
    pss_06 = SLB035B6,
    pss_07 = SLB035B7,  # reverse
    pss_08 = SLB035B8,  # reverse
    pss_09 = SLB035B9,
    pss_10 = SLB035B10,

    # Stressful Life Events (last 5 years)
    sle_01 = SLB037B1,
    sle_02 = SLB037B1A,
    sle_03 = SLB037B2,
    sle_04 = SLB037B2A,
    sle_05 = SLB037B3,
    sle_06 = SLB037B3A,
    sle_07 = SLB037B4,
    sle_08 = SLB037B4A,
    sle_09 = SLB037B5,
    sle_10 = SLB037B5A,
    sle_11 = SLB037B6,
    sle_12 = SLB037B6A,
    
  )

# Preview the cleaned 2022 data
glimpse(clean_2022)


```

### Optional: Save 2022-only dataset

```{r}
# Run this if you're only building 2022 and not doing longitudinal

write.csv(clean_2022, "HRS_Stress_2022.csv", row.names = FALSE)
saveRDS(clean_2022, "HRS_Stress_2022.rds")

```



---

### Combine All Years (Run Only for Longitudinal Build)

```{r}

# NOTE:
# - This step combines all the yearly "clean_[year]" datasets that you created above.
# - If you only built certain years (e.g., 2014–2022), remove any years from the
#   "clean_list" below that you did not load or clean.
#   Example:
#     clean_list <- list(clean_2014, clean_2016, clean_2018, clean_2020, clean_2022)
#
# The bind_rows() function will automatically align columns and fill in missing
# variables with NA where necessary.

library(dplyr)

# Collect all cleaned waves
clean_list <- list(
  clean_2006, clean_2008, clean_2010, clean_2012, clean_2014,
  clean_2016, clean_2018, clean_2020, clean_2022
)

# Join all years by stacking (fills missing cols with NA automatically)
full_hrs_stress <- bind_rows(clean_list) %>%
  arrange(id_year) %>%
  relocate(id_year, .before = 1) # move id_year to first column

# Preview the full data
glimpse(full_hrs_stress)

```


---

### Save and Export Longitudinal Dataset (2006–2022 Only)


Export the cleaned dataset for future use in your preferred format. Below are examples using both `write.csv()` and `saveRDS()`: 

```{r}

# Save longitudinal dataset (2006–2022)
write.csv(full_hrs_stress, "HRS_Stress_Longitudinal_2006_2022.csv", row.names = FALSE)
saveRDS(full_hrs_stress, "HRS_Stress_Longitudinal_2006_2022.rds")

```


---

### Optional: Merge your HRS Stress dataset with your main HRS Allostatic (over)Load dataset

If you’ve already built the **HRS Allostatic (over)Load dataset** (from HRS_dataset_construction.Rmd), you can merge it with your **HRS Stress** dataset — whether you built a *single-year* or *longitudinal* version.  

**Instructions:**  

1. Load your Allostatic dataset.  
2. Choose the appropriate stress dataset object:  
- `clean_[year]` → for single-year builds (e.g., `clean_2016`)  
- `full_hrs_stress` → for longitudinal builds (2006–2022)  
3. Merge using the shared key `id_year`.  

```{r}

# Load your Allostatic (over)Load dataset
# (use the same file format you saved earlier)
# hrs_al <- readRDS("hrs_allostatic_overload.rds")
# hrs_al <- read.csv("hrs_allostatic_overload.csv")

# ---- For single-year builds ----
# Replace [year] with your chosen year, e.g., clean_2016
# hrs_merged <- clean_[year] %>%
#   left_join(hrs_al, by = "id_year")

# ---- For longitudinal builds ----
# hrs_merged <- full_hrs_stress %>%
#   left_join(hrs_al, by = "id_year")

# Preview merged dataset
# glimpse(hrs_merged)

# Save/export merged dataset
# write.csv(hrs_merged, "HRS_Allostatic_Plus_Stress.csv", row.names = FALSE)
# saveRDS(hrs_merged, "HRS_Allostatic_Plus_Stress.rds")

```

