---
title: "HRS Geographical Dataset Construction"
subtitle: "CHORDS Lab – Washington State University"
author:
- "Jason Cross"
- "Shawna Beese"
date: "Last updated: 2025-08-30"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

## Overview

This R Markdown script walks through the construction of a supplementary geographic dataset for allostatic (over)load research using data from the [Health and Retirement Study (HRS)](https://hrs.isr.umich.edu/).

⚠️ Before continuing, make sure you’ve downloaded and the required raw data. See [`HRS_geo_access_guide.md`](https://insert_url_here) for instructions.

The [`HRS_geo_access_guide.md`](HRS_geo_access_guide.md) will walk you through downloading the data we will be pulling from in this script.
---

### What You’ll Need to Run This Script

- The unzipped folder `HRSXRegion2022v10Early/` (contains `HRSXREGION22.da` and `HRSXREGION22.sas`)
- R packages: SAScii, dplyr, stringr
- A target folder path to your `HRS Data Products` directory

## Setup & Libraries

```{r install-packages, eval=FALSE}
install.packages(c("SAScii", "dplyr", "stringr"))
```

```{r libraries}
library(SAScii)
library(dplyr)
library(stringr)
```

---

## 📁 Set File Path

```{r}

# Set to the location of your "HRS Data Products" folder
path <- "../path_to_your_folder/HRS Data Products/" # Replace with the full path to your local "HRS Data Products" folder

#####################
# Append Regional Data
#####################

# Set path to 2022 HRSXRegion folder
setwd(file.path(path, "HRSXRegion2022v10Early"))

region = read.SAScii("HRSXREGION22.da", "HRSXREGION22.sas")

full_data$HHID = unlist(lapply(full_data$unique_id,FUN = function(x){return(str_split_1(x[1],"_")[1])}))
region = region%>%mutate(unique_id = paste(HHID,PN,sep = "_"))

```

```{r}

# Select the required BEALE columns for 2003 and 2013 (from 2006-2022) -- include 2018, 2020 and 2022?
beale_columns <- c("BEALE2003_06", "BEALE2003_08", "BEALE2003_10", "BEALE2003_12", "BEALE2003_14", "BEALE2003_16", "BEALE2003_18", "BEALE2003_20", "BEALE2003_22",
                   "BEALE2013_06", "BEALE2013_08", "BEALE2013_10", "BEALE2013_12", "BEALE2013_14", "BEALE2013_16", "BEALE2013_18", "BEALE2013_20", "BEALE2013_22")

region_trim <- region %>%
  select(unique_id, all_of(beale_columns))  # Ensure `unique_id` is included with the BEALE columns

# Preview
glimpse(region_trim)

```

```{r}

geo_subset <- region_trim %>%
  # Pivot both Beale schemes across years
  pivot_longer(
    cols = matches("^BEALE(2003|2013)_(06|08|10|12|14|16|18|20|22)$"),
    names_to = c("scheme", "yr2"),
    names_pattern = "BEALE(2003|2013)_(\\d{2})",
    values_to = "value"
  ) %>%
  # Generate year names
  mutate(
    survey_year = 2000 + as.integer(yr2),         # 06 -> 2006, 22 -> 2022
    scheme = paste0("Beale_", scheme)
  ) %>%
  select(unique_id, survey_year, scheme, value) %>%
  # Pivot to get Beale_2003 and Beale_2013 columns side by side
  pivot_wider(
    names_from = scheme,
    values_from = value
  ) %>%
  # Build the id_year join key and select desired columns
  mutate(id_year = paste(unique_id, survey_year, sep = "_")) %>%
  select(id_year, unique_id, survey_year, Beale_2003, Beale_2013) %>%
  arrange(unique_id, survey_year) %>%

# Preview
glimpse(geo_subset)

```

---

### Save and Export 2006-2022 HRS Geographic Data

Export the cleaned dataset for future use in your preferred format. Below are examples using both `write.csv()` and `saveRDS()`: 

```{r}

# write.csv(geo_subset, "name_your_hrs_geo_dataset.csv")
# Example: write.csv(geo_subset, "hrs_geo_data.csv")

# saveRDS(geo_subset, "name_your_hrs_geo_dataset.rds")
# Example: saveRDS(geo_subset, "hrs_geo_data.rds")

```


---

### Optional: Merge your HRS Geographic dataset with your main HRS Allostatic (over)Load dataset

If you have already built the HRS Allostatic (over)Load dataset (from HRS/HRS_dataset_construction.Rmd), you can join it with geo_subset on the common key `id_year`.

```{r}

# Load your Allostatic (over)Load dataset
# (use the matching format you saved earlier)
# hrs_al <- readRDS("hrs_allostatic_overload.rds")
# hrs_al <- read.csv("hrs_allostatic_overload.csv")

# Join geo data with Allostatic (over)Load data on id_year
hrs_merged <- geo_subset %>%
  left_join(hrs_al, by = "id_year")

# Preview merged dataset
glimpse(hrs_merged)

# Save/export merged dataset
# write.csv(hrs_merged, "hrs_allostatic_plus_geo.csv", row.names = FALSE)
# saveRDS(hrs_merged, "hrs_allostatic_plus_geo.rds")

```

